---
title: "Homework 3"
author: Safiya Sirota
date: 2021-10-20
output: 
  github_document:
    toc: TRUE
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
  data("instacart")
  data("brfss_smart2010")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
nrow(instacart)
ncol(instacart)
names(instacart)
```

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(aisle_ct = n()) %>% 
  nrow()

instacart %>% 
  group_by(aisle) %>% 
  summarize(aisle_ct = n()) %>% 
  arrange(desc(aisle_ct)) %>% 
  top_n(5)
```

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(aisle_ct = n()) %>% 
  filter(aisle_ct >= 10000) %>% 
  ggplot(aes(x = fct_reorder(aisle, aisle_ct), y = aisle_ct, fill = aisle_ct)) +
  geom_col() +
  coord_flip()
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
baking <-
  instacart %>% 
    filter(aisle == "baking ingredients") %>% 
    group_by(product_name) %>% 
    summarize(product_ct = n()) %>% 
    arrange(desc(product_ct)) %>% 
    top_n(3)  %>% 
    mutate(
      aisle = "Baking Ingredients",
      rank = 1:3
    )

dog_food <-
  instacart %>% 
    filter(aisle == "dog food care") %>% 
    group_by(product_name) %>% 
    summarize(product_ct = n()) %>% 
    arrange(desc(product_ct)) %>% 
    top_n(3) %>% 
    mutate(
      aisle = "Dog Food & Care",
      rank = 1:3
    )

pkg_veg_fruit <-
  instacart %>% 
    filter(aisle == "packaged vegetables fruits") %>% 
    group_by(product_name) %>% 
    summarize(product_ct = n()) %>% 
    arrange(desc(product_ct)) %>% 
    top_n(3) %>% 
    mutate(
      aisle = "Packaged Vegetables & Fruits",
      rank = 1:3
    )

full_join(baking, dog_food) %>% 
  full_join(pkg_veg_fruit) %>% 
  select(aisle, rank, product_name, product_ct) %>% 
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour_of_day = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour_of_day
  ) %>% 
  rename(
      "Sunday" = `0`,
      "Monday" = `1`,
      "Tuesday" = `2`,
      "Wednesday" = `3`,
      "Thursday" = `4`,
      "Friday" = `5`,
      "Saturday" = `6`
      ) %>% 
  knitr::kable(digits = 1)

  
```

## Problem 2

Cleaning data

```{r}
brfss_smart2010_tidy <-
brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(
    topic == "Overall Health" & 
    response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")
    ) %>% 
  mutate(
    response = factor(response),
    response = forcats::fct_relevel(
      response, 
      c("Poor", "Fair", "Good", "Very good", "Excellent")
      )
  ) 
```

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_smart2010_tidy %>% 
  filter(year == 2002) %>% 
  select(locationabbr, locationdesc) %>% 
  group_by(locationabbr) %>% 
  summarize(county_ct = n_distinct(locationdesc)) %>% 
  filter(county_ct >= 7)

brfss_smart2010_tidy %>% 
  filter(year == 2010) %>% 
  select(locationabbr, locationdesc) %>% 
  group_by(locationabbr) %>% 
  summarize(county_ct = n_distinct(locationdesc)) %>% 
  filter(county_ct >= 7)
  
```

Construct a dataset that is limited to `Excellent` responses, and contains, year, state, and a variable that averages the `data_value` across locations within a state. Make a "spaghetti" plot of this average value over time within a state.

```{r}
brfss_smart2010_tidy %>% 
  filter(response == "Excellent") %>% 
  group_by(year, locationabbr) %>% 
  summarize(mean_data_value = mean(data_value, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_data_value, group = locationabbr)) +
  geom_line()
```

Make a two-panel plot showing, for the years 2006 and 2010, distribution of `data_value` responses ("Poor" to "Excellent") among locations in NY State.

```{r}
plot_2006 <-
  brfss_smart2010_tidy %>% 
  filter(year == 2006 & locationabbr == "NY") %>% 
  group_by(response) %>% 
  summarize(response_mean = mean(data_value)) %>% 
  ggplot(aes(x = response, y = response_mean)) +
  geom_col()

plot_2010 <-
  brfss_smart2010_tidy %>% 
  filter(year == 2010 & locationabbr == "NY") %>% 
  group_by(response) %>% 
  summarize(response_mean = mean(data_value)) %>% 
  ggplot(aes(x = response, y = response_mean)) +
  geom_col()

plot_2006 + plot_2010

```

## Problem 3